\section{Introduction}

Multirotor UAVs have been a topic of great research interest over the course of past decade. The first mathematical model of quadrotor vehicle capable of vertical takeoff and landing has been presented in \cite{hamel2002quad}. Since then, researchers have been working on various control designs with applications in payload transportation, formation flight, exploration, environment interaction etc.\\
In traditional UAV control methods the configuration is considered in Euclidean space as $\zeta = [x, y, z, \phi, \theta, \psi] \in \mathbb{R}^6$, where the first three components describe the UAV position in the inertial frame, while the latter three describe its orientation in the inertial frame \cite{uavModel}.

\noindent \textbf{Geometric Control}, on the other hand, considers the UAV configuration as a rigid body described by the location of its center of mass in $\mathbb{R}^3$ and attitude represented by the $3\times3$ orthonormal matrix. To discover the underlying structure of such a configuration geometric mechanics principles need to be applied. Originally introduced by Jurdjevic 1996. \cite{jurdjevic_1996}, Bloch 2004.\cite{bloch}, Bullo and Lewis \cite{bulloBook}, T. Lee 2008. \cite{Lee2008ComputationalGM}, T. Lee et al. 2018. \cite{LeeModel}, geometric mechanics is a modern description of classical mechanics from the perspective of differential geometry. \\
Therefore, it can now be stated that a set of $3\times3$ orthonormal matrices representing UAV attitude is a manifold, as there exists a local bijective map towards an $\mathbb{R}^3$ Euclidean space, i.e. each $3\times3$ matrix corresponds to a vector in $\mathbb{R}^3$ representing rotations around inertial frame axes. Assuming smooth maps and equipping the manifold with a group operator of matrix multiplication a Special orthogonal Lie group SO(3) consisting of $3\times3$ orthonormal matrices is subsequently defined. The final configuration is then a combination between translation in $\mathbb{R}^3$ and rotation in SO(3) resulting in a Special Euclidean Lie group SE(3) consisting of $4\times4$ transformation matrices containing both rotation and translation components. \\ 
Furthermore the UAV dynamics are, within this approach, considered as a Lagrangian or Hamiltonian system along with its unique properties: energy is conserved in absence of non-conservative forces; momentum map, a generalized notion of linear and angular momentum, with the system symmetry is preserved.

The main advantage of geometric control is the coordinate-free approach. Representing objects wrt. coordinate frames is often confusing and error prone. This approach only relies on compact expressions wrt. the inertial coordinate frame. To further elaborate, let the UAV attitude be taken into consideration. Expressed globally in the SO(3) configuration manifold no issues arise. If the parametrization to one of 24 types of Euler angles is applied, singularities may occur when multiple Euler angle sets correspond to a single SO(3) configuration. When expressing the attitude in quaternions sign ambiguities occur as there are always two possible quaternion solutions representing the SO(3) configuration \cite{euler}.

\noindent \textbf{Passive Decomposition} is the second control framework considered in this survey. The notion of energetic passivity of the closed-loop interactive mechanical system implies that the maximum energy that is able to be extracted from the system is always bounded. This proves to be a favorable feature in interactive mechanical systems such as: telemanipulators (e.g. a robot manipulated by a human operator), multi-agent collaborative manipulation (e.g. multiple robots carrying a payload) and human interactive robots (e.g. physical coupling with humans for direct assistance). Passivity based control has been a motivating factor for research by  Colgate 1994. \cite{passive6}, Colgate et al. 1997. \cite{passive12}, Stramigioli 2001. \cite{passive5}, P. Li 2001. \cite{passive8}. \\
The concept of passive decomposition was introduced by D. Lee 2008. \cite{LeePassive} and is formulated as follows. Let the dynamics of an interactive mechanical system be split into three components. Shape dynamics address the internal configuration of each robot along wrt. the holonomic constraints \cite{holonomic}. Locked dynamics represent the current overall behavior of multiple interactive robots with their internal configuration fixed. Finally, the coupled system dynamics come as a result of the interaction between locked and shape dynamics. It is, of course, implied that the underlying system configuration is a Lie group which evolves on its respective tangent bundle.

The main advantage of passive decomposition is that cancellation of the dynamic coupling does not require nor generate energy by itself. This enables decoupled control of shape and locked dynamics while enforcing energetic passivity. Furthermore, passivity of the closed-loop interactive mechanical system is guaranteed if the locked and shape system controls are designed to comply with energetic passivity condition for their respective systems.

This paper is organized as follows. In Section \ref{section:geometric_mechanics} a review of geometric control methods is given. In Section \ref{section:pssivity} the current accomplishments in passivity-based control methods is presented. It's worth noting that only applications for multi-rotor UAVs are taken into account. As for Section  \ref{future_work} a brief overview of novel applications and suggestions for advancements in current state-of-the-art is given. Section \ref{section:conclusion} gives a brief summary of the overall survey.